---
title: "Seurat_Code_Filtering_2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Install packages
```{r}
#install.packages("BiocManager")
#BiocManager::install("Seurat")
#BiocManager::install("ggplot2")
#install.packages("dplr")
#install.packages("ape")
#install.packages("cowplot")
#install.packages("Matrix")
#install.packages("clustree")
#install.packages("igraph")
install.packages('patchwork')
install.packages('stringr')
BiocManager::install('Nebulosa')
install.packages("qs")
# Installing scTransform v2
# install glmGamPoi
if (!requireNamespace("BiocManager", quietly = TRUE)) install.packages("BiocManager")
BiocManager::install("glmGamPoi")
# install sctransform from Github
devtools::install_github("satijalab/sctransform", ref = "develop")
devtools::install_github(repo = "samuel-marsh/scCustomize")

if (!require("devtools"))
  install.packages("devtools")
devtools::install_github("Winnie09/Palo")

```

Load packages into your workspace. 
```{r}
library(Seurat)
library(ggplot2)
library(dplyr)
library(ape)
library(cowplot)
library(Matrix)
library(sctransform)
library(clustree)
library(tidyverse)
library(magrittr)
library(RColorBrewer)
library(patchwork)
#library(glmGamPoi)
library(scCustomize)
library(qs)
library(Nebulosa)
library(Palo)

```


```{r}
#Specify your preferred directory for all input + output
dir= "/Users/phinguyen/Dropbox (Personal)/Hen Lab/Data/SingleCell RNA-seq/Seurat/Seurat_HN5-8_Multi_Joint/1_SCT_2022-9-6.RData"

#Specify exactly where your seurat files are
# datafolder = "Seurat_HN1-4"

#the EXACT data file you want to open from the file you specified with "datafolder"
# filename = "1_DG_SCT_2023-1-17.RData"

#set up folders
QCfolder = "QC"
Plotfolder = "Plots"

#This name needs to match your project name within the seurat object
#project<-"AgingSC" 

#Specify your organism; capitalize the first letter (i.e. "Mouse", "Human","Zebrafish","Rat")
organism = "Mouse"

#metrics you want to look at for QC
m = c("nCount_RNA","nFeature_RNA","percent.mito")

#You can add in housekeeping genes if you want them to be scaled always; otherwise set to "NULL"
add.genes = NULL

#metadata dimensions you want to cluster on
#dims = c("seurat_clusters","ID","Cell_Type","Condition")

#edit to include all desired celltypes to subset on later; if not subsetting, set to "all"
keep = "all"

```


Load in your filtered dataset
```{r}
# load(file.path(dir,datafolder,filename))
load(file.path(dir))
```

Function to print multiple graphs: 
```{r}
PrintSeuratGraph = function(namecard = "a",seurat_object = sobject,graphtype = "feature",feature = NULL,group = NULL,split=NULL,cellnames=NULL){
  if (!is.null(cellnames)){
    Idents(seurat_object) = cellnames[1]
    cells = colnames(seurat_object)[Idents(seurat_object) %in% cellnames[2:length(cellnames)]]} 
  else {cells = cellnames}
  if (graphtype == "feature"){
    graph = FeaturePlot(seurat_object,features = feature,split.by = split, cells = cells,cols = c("lightyellow","darkred"))
  }
  if (graphtype == "violin"){
    graph = VlnPlot(seurat_object,features = feature, pt.size = 0.1, idents = cellnames[2:length(cellnames)],group.by = group, split.by = split)
  }
  if (graphtype == "dim"){
    graph = DimPlot(seurat_object,cells = cells, group.by = group, split.by = split)
    
  }
  
  name = paste0(feature,"_",graphtype,namecard,".eps")
  graph
  setEPS()
  postscript(file.path(dir,Plotfolder,name))
  print(graph) 
  dev.off()
}

```



*##############################**##############################*

#Find variable features, normalize, scale, run PCA, clustering, umap
*SCTransform*
Normalization and variance stabilization (Hafemeister and Satija, 2019 https://pubmed.ncbi.nlm.nih.gov/31870423/)
```{r, results='hide',eval = FALSE}
# Good idea to save prior to this step b/c may crash R.

#Choose what to regress out for sctransform- could be age, sex, or any metadata column
regress = c("nCount_RNA", "percent.mito") #"orig.ident"

sobject <- SCTransform(sobject, vars.to.regress = regress, assay = "RNA", verbose = T, variable.features.n = 3000, conserve.memory = T, vst.flavor = NULL)
                         #vst.flavor = "v2" #vst.flavor v2 specifies SCtransform version 2
                        # method = "glmGamPoi"
# "limit reachediteration" error message is ok.

```

Show most variable genes
```{r}
labels <- c(head(VariableFeatures(sobject),20),add.genes)
plot1 = VariableFeaturePlot(sobject)
LabelPoints(plot=plot1, points = labels, repel = F, xnudge = 0.1, ynudge = 0.5)
```

*##############################**##############################*

*Run PCA analysis and show elbow plot*
```{r}
sobject <- RunPCA(sobject,features = VariableFeatures(sobject),npcs = 50, verbose = FALSE)
ElbowPlot(sobject,ndims = 50, reduction = "pca")


print(sobject[["pca"]], dims = 1:20, nfeatures = 15)
# Output will be 5 genes per PC. PCs enriched with Mito/heatshock/ribo genes can be excluded

# Plot PCA
PCAPlot(sobject, split.by = "condition")  
PCAPlot(sobject, group.by = "condition")  

#scCustomize heatmap of defined PC
PC_Plotting(seurat_object = sobject, dim_number = 1)

```


Which PC's to include for downstream clustering? Can estimate qualitatively based on elbow plot or take quantitative approach to find PC's in which: 
1.) The point where the PCs only contribute 5% of standard deviation and the principal components cumulatively contribute 90% of the standard deviation.
2.) The point where the percent change in variation between the consecutive PCs is less than 0.1%.
```{r}
# Find the first metric (1)
# Determine percent of variation associated with each PC
pct <- sobject[["pca"]]@stdev / sum(sobject[["pca"]]@stdev) * 100

# Calculate cumulative percents for each PC
cumu <- cumsum(pct)

# Determine which PC exhibits cumulative percent greater than 90% and % variation associated with the PC as less than 5
metric1 <- which(cumu > 90 & pct < 5)[1]
print("Last PC where cumulative % of SD is 90%")
metric1

# Find the second metric (2)
# Determine the difference between variation of PC and subsequent PC
metric2 <- sort(which((pct[1:length(pct) - 1] - pct[2:length(pct)]) > 0.1), decreasing = T)[1] + 1
print("Last PC where % change of variation is more than 0.1%.")
metric2

```

```{r}
# PCs for dimensionality reduction. Can change if one of the pc's is for something you think is a technical error (i.e. HSP, RP, etc)
pcs = c(1:30)

print(sobject[["pca"]], dims = 1:20, nfeatures = 8)
```

Once you are satisfied with pc's, run clustering (UMAP, FindNeighbors, FindClusters): 
```{r}
# number of PC's defined at top of chunk above
sobject<-RunUMAP(sobject,reduction = "pca",dims = pcs, verbose = F, n.neighbors = 30, min.dist = 0.3, n.epochs = NULL) 
# Defaults for n.neighbors = 30 and min.dist = 0.3

# Determine the K-nearest neighbor graph
sobject<-FindNeighbors(sobject,dims=pcs,verbose=F) 

# Quick UMAP test
Idents(sobject) = "condition"
DimPlot(sobject, label = T, label.size = 4, pt.size = .5, 
        label.color = "black", order = F, repel = F)

```



Determine the clusters for various resolutions
```{r}
# clustering resolution; the last number will be saved as "seurat_clusters" in metadata
res = c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9)
#sobject<-FindClusters(sobject,verbose=F,resolution = res)
sobject<-FindClusters(sobject,verbose=F,resolution = 0.9)

#View(sobject@meta.data)
#head(sobject@meta.data)


# Test how UMAP plot looks with different clustering resolutions

Idents(sobject) = "seurat_clusters"
for (Cluster_resolution in c("SCT_snn_res.0.1",
                             "SCT_snn_res.0.2",
                             "SCT_snn_res.0.3",
                             "SCT_snn_res.0.4",
                             "SCT_snn_res.0.5",
                             "SCT_snn_res.0.6",
                             "SCT_snn_res.0.7",
                             "SCT_snn_res.0.8",
                             "SCT_snn_res.0.9")){
  Idents(sobject) = Cluster_resolution
  print(DimPlot(sobject, label = T, label.size = 6, pt.size = 1.0) + ggplot2::ggtitle(label = Cluster_resolution))
      }

```



Clustree package to generate clustering trees- see how cells/clusters move as clustering resolution increases. Vignette: https://cran.r-project.org/web/packages/clustree/vignettes/clustree.html 
```{r}
clustree(sobject, prefix = "SCT_snn_res.") + theme(legend.position = "bottom")

# Alternate algorithm for aesthetic variety
#pdf(file.path(dir,"Plots", paste0("testClustree.pdf")))
    clustree(sobject, prefix = "SCT_snn_res.", layout = "sugiyama", use_core_edges = FALSE) + 
    guides(edge_colour = FALSE, edge_alpha = FALSE) + theme(legend.position = "bottom")
#dev.off()

# Overlay expression of gene of interest
clustree(sobject, prefix = "SCT_snn_res.", layout = "sugiyama", use_core_edges = FALSE, node_colour = "Prox1", node_colour_aggr = "mean") + guides(edge_colour = FALSE, edge_alpha = FALSE) + theme(legend.position = "bottom") + scale_color_gradient(low = "lightblue", high = "red")
    
#clustree(sobject, prefix = "SCT_snn_res.", 
        # node_colour = "nFeature_SCT", node_colour_aggr = "mean", 
        # node_label = "nFeature_SCT", node_label_aggr = "max")

#clustree(sobject, prefix = "SCT_snn_res.", 
        # node_colour = "sc3_stability") #higher sc3_stability, the more stable the cluster across resolutions

# Opacity of arrows indicates relative contribution of cells moving towards a cluster of a subsequent clustering resolution
```


*##############################**##############################*

*PCA plots*
```{r}
Idents(sobject) = "seurat_clusters"

# PCA
DimPlot(sobject, reduction = "pca", group.by = "condition", dims = c(1,2)) # dims sets the PCs used for display
DimPlot(sobject, reduction = "pca", group.by = "condition", dims = c(3,4))
DimPlot(sobject, reduction = "pca", split.by = "condition", dims = c(1,2))
#DimPlot_scCustom(sobject, reduction = "pca", split.by = "condition", dims = c(3,4), num_columns = 4, label = F)

#setEPS()
#postscript(file.path(dir,"Plots", paste0("name.eps")))
#DimPlot()
#dev.off()
```

*palopal* setup for optimized color palette
```{r}
UMAP <- sobject[["umap"]]@cell.embeddings

CellClusters <- Idents(sobject)

# Generate a color palette which is used by ggplot2
gg_color_hue <- function(n) {
  hues = seq(30, 500, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]
}
pal <- gg_color_hue(length(unique(CellClusters)))

# Run Palo to get the optimized palette
palopal <- Palo(UMAP,CellClusters,pal)
palopal # can use as input into cols argument of DimPlot
```

Feature and violin plots- QC metrics. 
```{r}
Idents(sobject) = "seurat_clusters"

for (QC_metrics in c(m, "percent.ribo")){
  Idents(sobject) = "seurat_clusters"
  print(FeaturePlot(sobject, feature=QC_metrics,
            pt.size = 0.4, label = T, label.size = 5, order = T, 
            ))
  Idents(sobject) = "ID"
  print(VlnPlot(sobject, feature=QC_metrics, pt.size = 0))
}

# Feature and Violin Plots of QC metrics (split by ID)
for (QC_metrics in m){
  #print(FeaturePlot(sobject, feature=QC_metrics,
           # pt.size = 0.4, label = T, label.size = 5, order = T, 
           # split.by = "ID"))
  print(VlnPlot(sobject, feature=QC_metrics, split.by = "condition", group.by = "seurat_clusters", pt.size = 0.1))
}

table(sobject$ID, sobject$seurat_clusters)
```



*Clustering Tree*
```{r}
#Build a clustering tree
Idents(sobject) = "seurat_clusters"
pcs = c(1:30)

sobject= BuildClusterTree(sobject,dims = pcs)
tree = sobject@tools$BuildClusterTree
plot.phylo(tree, use.edge.length = T, direction = "downwards", 
           node.pos = 1, srt = 90, adj = 0.5, label.offset = 0.75, font = 1, cex = 1.75, no.margin = T)
```




*##############################**##############################*

Save QC Violin or Feature plots
```{r}
setEPS()
postscript(file.path(dir, "Plots", paste0("2_VlnPlot_4-7-21.eps")))
VlnPlot(sobject, feature = "Hexb", pt.size = 0.00)
dev.off()
```



Exporting as EPS or pdf
```{r}
#Export EPS files of
setEPS()
  postscript(file.path(dir,QCfolder,paste0(filename,"_percent.mito.eps")))
  # Can put feature/violin plot command below
  VlnPlot(sobject,"percent.mito", pt.size = 0.0) 
  #
  dev.off()
  
# Copying table information
table_to_export = table(sobject$ID, sobject$seurat_clusters)
table_to_export
# change table into data frame, which you can then copy + paste special into excel
table_to_export2 <- as.data.frame.matrix(table_to_export)
table_to_export2
table(sobject$ID)

table(sobject$ID, sobject$seurat_clusters, sobject$sex)
```


*##############################**##############################*

Remove a metadata column
```{r}
# To remove a metadata column (either of these works)

# sobject$cluster_age <- NULL
# sobject[['SCT_snn_res.0.55']] <- NULL
```

Create new metadata columns
```{r}
#For a given cluster that appears to segregate by sample/condition, you can create new metadata column to separate cluster 1 old vs cluster 1 young and compare with FindMarkers to see the differences
sobject$cluster_age = paste0(sobject$seurat_clusters,"_",sobject$age)
Idents(sobject) = "cluster_age"
#Idents(sobject) = "seurat_clusters"

sobject@meta.data %>% 
         View()

DimPlot(sobject,
        split.by = "age", label.size = 4, pt.size = 0.6,
        label = T) 
```



Changing the order of values in a level
```{r}
#If you don't have Veh listed first (for ex. with the table function), then you can reorder them

sobject$condition <- factor(sobject$condition, levels = c("Veh","Cort","Flx","ECS"))

```


*##############################**##############################*

```{r}
dir = "/Users/phinguyen/Dropbox (Personal)/Hen Lab/Data/SingleCell RNA-seq/Seurat/Seurat_HN1-4"

save(sobject,file = file.path(dir,paste0("1_DG_SCT_2023-1-17.RData")))
```



*##############################**##############################*


****