---
title: "Untitled"
output: html_document
date: "2023-04-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(Seurat)
library(ggplot2)
library(dplyr)
library(ape)
library(cowplot)
library(Matrix)
library(sctransform)
library(clustree)
library(tidyverse)
library(magrittr)
library(RColorBrewer)
library(patchwork)
#library(glmGamPoi)
library(scCustomize)
library(qs)
library(Nebulosa)
library(Palo)
library(viridis)
library(Hmisc) # Cs() command adds quotes around each gene in a comma-separated gene list
```


```{r}
dir = "/Users/phinguyen/Library/CloudStorage/Dropbox/Hen Lab/Data/SingleCell RNA-seq/1_Neurons-Microglia_Veh-Cort-Flx-ECS_May 2022/Seurat/Seurat_HN1-4/1_DG_SCT_2023-9-11_Dematured Analysis_Res 1-1"

load(file.path(dir))
```



******************************************************
*UMAP*

*palopal* setup for optimized color palette
```{r}
Idents(sobject) = "seurat_clusters"

UMAP <- sobject[["umap"]]@cell.embeddings

CellClusters <- Idents(sobject)

# Generate a color palette which is used by ggplot2
gg_color_hue <- function(n) {
  hues = seq(30, 500, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]
}
pal <- gg_color_hue(length(unique(CellClusters)))

# Run Palo to get the optimized palette
palopal <- Palo(UMAP,CellClusters,pal)
palopal # can use as input into cols argument of DimPlot
```


*UMAP - by CLUSTERS*
```{r}
Idents(sobject) = "seurat_clusters"
label = T
repel = F
label.size = 5
pt.size = 0.5
raster = F

DimPlot(sobject, label = label, repel = repel, label.size = label.size, pt.size = pt.size, order = T,
        label.color = c("black"),
        cols = palopal,
        raster = raster) +
              theme(aspect.ratio=1, # sets square plot
              axis.title.x = element_text(hjust=0), axis.title.y = element_text(hjust=0), # adjust axes labels to corner
              axis.line = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), # hides axes and ticks
              panel.border = element_rect(colour = "black", fill=NA, size=1)) + # add border
              labs(y= "UMAP 2", x = "UMAP 1") 
```


*UMAP - by cell_type*
```{r}
Idents(sobject) = "cell_type"
label = F
repel = F
label.size = 5
pt.size = 1.5
raster = T


DimPlot(sobject, label = label, repel = repel, label.size = label.size, pt.size = pt.size, order = T,
        label.color = c("black"),
        raster = raster) +
              theme(aspect.ratio=1, # sets square plot
              axis.title.x = element_text(hjust=0), axis.title.y = element_text(hjust=0), # adjust axes labels to corner
              axis.line = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), # hides axes and ticks
              panel.border = element_rect(colour = "black", fill=NA, size=1)) + # add border
              labs(y= "UMAP 2", x = "UMAP 1") 

```


*UMAP - by condition*
```{r}
Idents(sobject) = "condition"
label = F
repel = F
label.size = 5
pt.size = 0.5
raster = F


DimPlot(sobject, label = label, repel = repel, label.size = label.size, pt.size = pt.size, order = c("Veh","Cort","Flx","ECS"),
        label.color = c("black"),
        cols = c("seagreen3","slateblue3","darkorange1","black"),
        raster = raster) +
              theme(aspect.ratio=1, # sets square plot
              axis.title.x = element_text(hjust=0), axis.title.y = element_text(hjust=0), # adjust axes labels to corner
              axis.line = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), # hides axes and ticks
              panel.border = element_rect(colour = "black", fill=NA, size=1)) + # add border
              labs(y= "UMAP 2", x = "UMAP 1") 
```



UMAP for each condition (showing cells from condition over non-condition)
```{r}
Idents(sobject) = "condition"
levels(sobject)

pt.size = 0.5

Cluster_Highlight_Plot(sobject, cluster_name = "Veh", highlight_color = "black", pt.size = pt.size,
    background_color = "lightgray") + ggplot2::ggtitle(label = "Veh") +
              theme(aspect.ratio=1, # sets square plot
              axis.title.x = element_text(hjust=0), axis.title.y = element_text(hjust=0), # adjust axes labels to corner
              axis.line = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), # hides axes and ticks
              panel.border = element_rect(colour = "black", fill=NA, size=1)) + # add border
              labs(y= "UMAP 2", x = "UMAP 1") 

Cluster_Highlight_Plot(sobject, cluster_name = "Cort", highlight_color = "darkorange1", pt.size = pt.size,
    background_color = "lightgray") + ggplot2::ggtitle(label = "Cort") +
              theme(aspect.ratio=1, # sets square plot
              axis.title.x = element_text(hjust=0), axis.title.y = element_text(hjust=0), # adjust axes labels to corner
              axis.line = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), # hides axes and ticks
              panel.border = element_rect(colour = "black", fill=NA, size=1)) + # add border
              labs(y= "UMAP 2", x = "UMAP 1") 

Cluster_Highlight_Plot(sobject, cluster_name = "Flx", highlight_color = "slateblue3", pt.size = pt.size,
    background_color = "lightgray") + ggplot2::ggtitle(label = "Flx") +
              theme(aspect.ratio=1, # sets square plot
              axis.title.x = element_text(hjust=0), axis.title.y = element_text(hjust=0), # adjust axes labels to corner
              axis.line = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), # hides axes and ticks
              panel.border = element_rect(colour = "black", fill=NA, size=1)) + # add border
              labs(y= "UMAP 2", x = "UMAP 1") 

Cluster_Highlight_Plot(sobject, cluster_name = "ECS", highlight_color = "seagreen3", pt.size = pt.size,
    background_color = "lightgray") + ggplot2::ggtitle(label = "ECS") +
              theme(aspect.ratio=1, # sets square plot
              axis.title.x = element_text(hjust=0), axis.title.y = element_text(hjust=0), # adjust axes labels to corner
              axis.line = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), # hides axes and ticks
              panel.border = element_rect(colour = "black", fill=NA, size=1)) + # add border
              labs(y= "UMAP 2", x = "UMAP 1") 

```


```{r}
Idents(sobject) = "cell_type"
levels(sobject)

pt.size = 0.25

Cluster_Highlight_Plot(sobject, cluster_name = "CA3", highlight_color = "black", pt.size = pt.size,
    background_color = "lightgray") + ggplot2::ggtitle(label = "Cort") +
              theme(aspect.ratio=1, # sets square plot
              axis.title.x = element_text(hjust=0), axis.title.y = element_text(hjust=0), # adjust axes labels to corner
              axis.line = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), # hides axes and ticks
              panel.border = element_rect(colour = "black", fill=NA, size=1)) + # add border
              labs(y= "UMAP 2", x = "UMAP 1") 

Cluster_Highlight_Plot(sobject, cluster_name = "CA1", highlight_color = "brown2", pt.size = pt.size,
    background_color = "lightgray") + ggplot2::ggtitle(label = "Flx_5d") +
              theme(aspect.ratio=1, # sets square plot
              axis.title.x = element_text(hjust=0), axis.title.y = element_text(hjust=0), # adjust axes labels to corner
              axis.line = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), # hides axes and ticks
              panel.border = element_rect(colour = "black", fill=NA, size=1)) + # add border
              labs(y= "UMAP 2", x = "UMAP 1") 

Cluster_Highlight_Plot(sobject, cluster_name = "Oligodendrocytes", highlight_color = "cornflowerblue", pt.size = pt.size,
    background_color = "lightgray") + ggplot2::ggtitle(label = "Flx_3wks") +
              theme(aspect.ratio=1, # sets square plot
              axis.title.x = element_text(hjust=0), axis.title.y = element_text(hjust=0), # adjust axes labels to corner
              axis.line = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), # hides axes and ticks
              panel.border = element_rect(colour = "black", fill=NA, size=1)) + # add border
              labs(y= "UMAP 2", x = "UMAP 1") 

Cluster_Highlight_Plot(sobject, cluster_name = "OPCs", highlight_color = "forestgreen", pt.size = pt.size,
    background_color = "lightgray") + ggplot2::ggtitle(label = "Flx_3wks") +
              theme(aspect.ratio=1, # sets square plot
              axis.title.x = element_text(hjust=0), axis.title.y = element_text(hjust=0), # adjust axes labels to corner
              axis.line = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), # hides axes and ticks
              panel.border = element_rect(colour = "black", fill=NA, size=1)) + # add border
              labs(y= "UMAP 2", x = "UMAP 1") 
Cluster_Highlight_Plot(sobject, cluster_name = "CA2", highlight_color = "brown2", pt.size = pt.size,
    background_color = "lightgray") + ggplot2::ggtitle(label = "Flx_3wks") +
              theme(aspect.ratio=1, # sets square plot
              axis.title.x = element_text(hjust=0), axis.title.y = element_text(hjust=0), # adjust axes labels to corner
              axis.line = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), # hides axes and ticks
              panel.border = element_rect(colour = "black", fill=NA, size=1)) + # add border
              labs(y= "UMAP 2", x = "UMAP 1") 
Cluster_Highlight_Plot(sobject, cluster_name = "Mossy", highlight_color = "brown2", pt.size = pt.size,
    background_color = "lightgray") + ggplot2::ggtitle(label = "Flx_3wks") +
              theme(aspect.ratio=1, # sets square plot
              axis.title.x = element_text(hjust=0), axis.title.y = element_text(hjust=0), # adjust axes labels to corner
              axis.line = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), # hides axes and ticks
              panel.border = element_rect(colour = "black", fill=NA, size=1)) + # add border
              labs(y= "UMAP 2", x = "UMAP 1") 
```




******************************************************
*Feature Plots*


iGenes
```{r}
Neurons = c("Rbfox3","Slc17a7","Syn1","Hexb","Gfap","Mbp") #Slc17a7 is vglut1

DG = c("Prox1","Stxbp6","Dock10","Pde7b","Calb1","Pomc","Rbfox3")
Immature = c("Prox1","Dcx","Stmn1","Ascl1","Igfbpl1","Neurod1","Sox11","Cd24a","Tmsb10", "Mki67","Calb1","Ablim3","Aldoc","Sox9","Hes5","Gfap")

DGvsPyr = c("Dkk3","Ociad2","Cpne7") # non-DGGCs is Dkk3 and pyramids are Ociad2
dDG_vDG = c("Prox1","Pdzd2","Lct","Gsg1l","Spata13","Stra6","Trhr","Cpne7","Grp","Nr2f2") #dDG up until Trhr
dHipp = c("Cadm2","Epha7","Epha5")
vHipp = c("Cadm1","Nrbp2")

CA1 = c("Ccn3","Mpped1","Fibcd1","Satb2", "Bcl11b")
CA2 = c("Cacng5","Tgfb1i1","Srl","Pygo1","F2r","Dusp5","Fgf5","Rgs14","Amigo2","Pcp4")
CA3 = c("Prkcd","Vgll3","Coch","Grik4","Elavl2",
        "Csf2rb2") #Csf2rb2 for MC and vCA3

Mossy = c("Calb2","Csf2rb2","Nmb","Ap2s1","Ass1","Ajap1","Dusp26","Rgs12","Fgf1","Serpinf1","Emid1","Cda",
       "Glipr1","Necab1","Calb2","Drd2","Sv2b","Foxg1") #Csf2rb2 for MC and vCA3, Foxg1 from Hochgerner 2018

GABA = c("Gad1","Gad2","Slc32a1","Slc6a1","Erbb4","Pvalb","Sst","Cck","Vip","Calb2")
Cajal_Retzius = c("Reln","Lhx5","Nhlh2","Trp73", "Diablo","Ndnf","Cacna2d2")


Glia = c("Hexb","Ifitm3","Gfap","Apoe","Aqp4","Nes","Mki67","Mbp","Olig2","Cspg4","Col4a1")
Microglia = Cs(Hexb,Cx3cr1,P2ry12,Aif1,Tmem119,Trem2,Gfap,Mbp,Rbfox3,Lyz2,Mrc1,Lyve1) #Lyz2=peripheral m0, Mrc1/Lyve1=perivascular m0, from Zeisel 2015
Astrocytes = Cs(Aldh1l1,Gfap,Aqp4,Serpinf1,Gja1,Mfge8,Slco1c1,Sox9)
Oligos = Cs(Trem2,Hexb,Gfap,Aqp4,Mbp,Mog,Cspg4,Pdgfra,Itpr2,Cnksr3,Rhob,Omg,Klk6) # Cspg4=Ng2,Pdgfra/NG2=OPC, Itpr2/Cnksr3=early oligo, Rhob/Omg=mid oligo, Klk6=late oligo -> from Zeisel 2015 
Vascular = Cs(Nes,Prom1,Cldn5,Cdh5,Pecam1,Ocln,Flt1,Esam,Abcb1a,Vwf,Mfsd2a,Vtn,Higd1b)
Pericyte = Cs(Cspg4,Pdgfrb,Anpep,Rgs5,Kcnj8,Abcc9,S1pr3,Mcam,Ifitm1,Baiap3,Ehd3)


IEG = c("Fos","Arc","Npas4","Egr1","Penk","Acan")
Opioid = c("Oprd1", "Oprm1","Penk", "Necab3", "Col6a1", "Htr1a","Htr2a","Htr4","Bdnf","Sv2a","Sv2b","Sv2c","Homer1") 
Serotonin = c("Htr1a","Htr1b","Htr2a","Htr2b","Htr3a","Htr4","Htr5a","Htr7","Gabbr1","Gabbr1","Gria1","Gria2")

```


*FeaturePlots* of iGenes (FeaturePlot)
```{r}
Idents(sobject) = "condition"

igenes = DG

min.cutoff = 0
label.size = 5
pt.size = 0.1

pal <- viridis(n = 10, option = "C", direction = 1)

for (igene_variable in igenes){
 print(FeaturePlot(sobject, features = igene_variable, 
          pt.size = pt.size, label = F, repel = T, label.size = label.size,
          cols = pal,
          order = T, 
          min.cutoff = min.cutoff,
          raster = F) + 
              theme(aspect.ratio=1, # sets square plot
              axis.title.x = element_text(hjust=0), axis.title.y = element_text(hjust=0), # adjust axes labels to corner
              axis.line = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), # hides axes and ticks
              panel.border = element_rect(colour = "black", fill=NA, size=1)) + # add border
              labs(y= "UMAP 2", x = "UMAP 1") + DarkTheme() 
          )}
```


******************************************************

*DotPlots*
```{r}
dotplot_genelist = Cs(Gfap,Satb2,Amigo2,Grik4,Prox1,Igfbpl1,Pecam1,Gad2,Lyve1,Tmem119,Drd2,Mbp,Pdgfra)


dot.scale = 10 # max size of the dot
scale.min = 0 #lower limit of % expressed
scale.max = 100 #upper limit of % expressed
col.min = 0.0 #lower limit of avg expression
col.max = 3 #upper limit of avg expression
aspect.ratio = 0.4 # sets square plot


DotPlot(sobject, features = dotplot_genelist,
        col.min = col.min,
        col.max = col.max,
        dot.scale = dot.scale, 
        scale.min = scale.min, scale.max = scale.max) + 
              theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),
              aspect.ratio = aspect.ratio,
              axis.line = element_blank(), # hides axes and ticks
              panel.border = element_rect(colour = "black", fill=NA, size=1)) + 
                xlab("") + ylab("") + 
                coord_flip() +
                scale_colour_gradient2(high = "brown1")

```


*Module scores*

Hongjunsong gene list top50
```{r}
Song_Immature <- Cs(
Igfbpl1,
Neurod1,
Sox11,
Cd24a,
Tmsb10,
Prox1,
Cplx2,
Nfix,
Zbtb20,
Bcl11a,
Stmn1,
Ctxn1,
Tubb2b,
Cnih2,
Mllt11,
Pcp4,
Dcx,
Nfib,
Map1b,
Hist3h2ba,
Marcksl1,
Tuba1a,
Tagln3,
Tubb3,
Epha4,
Serf1,
Tubb5,
Pak7,
H3f3b,
Nsg1,
Prdx2,
Thra,
Nnat,
Celf2,
Sstr2,
Rbfox1,
Synpr,
Nfia,
Rtn1,
Ptma,
Sema3c,
Snca,
Rbfox2,
Elavl3,
Cirbp,
Eif3e,
Bhlhe22,
Tcf4,
Gng2,
Auts2,
Ppp1r14c,
Eef1g,
Stmn4,
Stmn2
)
```

Module scores - Dot Plot
```{r}
# module scores
modules = Cs(Stmn1,Igfbpl1,Dcx,Song_Immature_1)
  #c("P132_1","P18_1","P5_1")
  #"abGC_score_1"
  #c("Granule_Immature_1","Neuroblast2_1")
  #"Cs(Stmn1,Igfbpl1,Dcx,Song_Immature_1)
# Cs(Igfbpl1,P5_1,Dock10,P132_1)


dot.scale = 6.5 # max size of the dot
scale.min = 0 #lower limit of % expressed
scale.max = 50 #upper limit of % expressed
col.min = 0.1 #lower limit of avg expression
col.max = 0.75 #upper limit of avg expression
aspect.ratio = 1.8 # sets square plot


DotPlot(sobject, features = modules,
        col.min = col.min,
        col.max = col.max,
        dot.scale = dot.scale, 
        scale.min = scale.min, scale.max = scale.max) + 
              theme(axis.text.x = element_text(angle = 0, vjust = 0, hjust=0.5, size = 15), axis.text.y = element_text(size=15),
              aspect.ratio = aspect.ratio,
              axis.line = element_blank(), # hides axes and ticks
              panel.border = element_rect(colour = "black", fill=NA, size=1)) + 
                xlab("") + ylab("") + 
                coord_flip() +
                scale_colour_gradient2(high = "brown1")

```


Module scores - Feature Plot
```{r}
# FeaturePlot

min.cutoff = 0
na_cutoff = 0.0

for (igene_variable in modules){
 print(FeaturePlot_scCustom(sobject, features = igene_variable, 
          pt.size = pt.size, label = T, repel = T, label.size = label.size,
          
          order = T, 
          min.cutoff = min.cutoff, na_cutoff = na_cutoff) + 
              theme(aspect.ratio=1, # sets square plot
              axis.title.x = element_text(hjust=0), axis.title.y = element_text(hjust=0), # adjust axes labels to corner
              axis.line = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), # hides axes and ticks
              panel.border = element_rect(colour = "black", fill=NA, size=1)) + # add border
              labs(y= "UMAP 2", x = "UMAP 1") 
          )}
```





******************************************************
*Violin Plots*

```{r}
vlnplot_genelist = Cs(Calb1,Dcx,Igfbpl1,Stmn1)
```

```{r}
#Idents(sobject) = "ID"
levels(sobject)
pt.size = 0.0

VlnPlot(sobject, feature = vlnplot_genelist, pt.size = pt.size, slot = 'scale.data') + 
    xlab("")

#VlnPlot(sobject, feature = vlnplot_genelist, pt.size = pt.size, slot = 'counts') + 
    xlab("")

#VlnPlot(sobject, feature = vlnplot_genelist, pt.size = pt.size) + 
    xlab("")

```


*##############################**##############################*

Feature and violin plots- QC metrics
```{r}
Idents(sobject) = "seurat_clusters"

m = c("nCount_RNA","nFeature_RNA","percent.mito")

for (QC_metrics in c(m, "percent.ribo")){
  Idents(sobject) = "seurat_clusters"
  print(FeaturePlot(sobject, feature=QC_metrics,
            pt.size = 0.4, label = T, label.size = 5, order = T, 
            ))
  Idents(sobject) = "seurat_clusters"
  print(VlnPlot(sobject, feature=QC_metrics, pt.size = 0))
}

# Feature and Violin Plots of QC metrics (split by ID)
for (QC_metrics in m){
  #print(FeaturePlot(sobject, feature=QC_metrics,
           # pt.size = 0.4, label = T, label.size = 5, order = T, 
           # split.by = "ID"))
  print(VlnPlot(sobject, feature=QC_metrics, split.by = "condition", group.by = "seurat_clusters", pt.size = 0.1))
}

table(sobject$Mouse_ID, sobject$seurat_clusters)
```


*##############################**##############################*

*Clustering Tree*
```{r}
#Build a clustering tree
Idents(sobject) = "seurat_clusters"
pcs = c(1:30)

sobject= BuildClusterTree(sobject,dims = pcs)
tree = sobject@tools$BuildClusterTree
plot.phylo(tree, use.edge.length = T, direction = "downwards", 
           node.pos = 1, srt = 90, adj = 0.5, label.offset = 0.75, font = 1, cex = 1.75, no.margin = T)
```



*##############################**##############################*


```{r}
dir = "/Users/phinguyen/Library/CloudStorage/Dropbox/Hen Lab/Data/SingleCell RNA-seq/2_AllCells_Early-Late-Flx_June 2023/Seurat/Run 2"

save(sobject,file = file.path(dir,paste0("2_all cells_COMBINED POST-FILTERING_no C5_SCT_2023-11-28.RData")))
```











